---
title: "Model-based Dimensionality Reduction for Single-cell RNA-seq with Generalized Bilinear Models"
author: "R package version 0.1.0"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation 

From the R console, `devtools::install_github("phillipnicol/scGBM")`. 

## Demo 

```{r}
library(scGBM)
library(ggplot2)
set.seed(1126490984)
```

Generate a count matrix that is random Poisson noise:


```{r, echo}
I <- 500
J <- 500
Y <- matrix(rpois(I*J,lambda=1),nrow=I,ncol=J)
colnames(Y) <- 1:J; rownames(Y) <- 1:I
```


Run scGBM with $M = 10$ latent factors

```{r}
out <- gbm.sc(Y,M=10)
```

Cluster the cell scores using Seurat

```{r, echo=T, results="hide"}
library(Seurat)
Sco <- CreateSeuratObject(counts=Y)
colnames(out$V) <- 1:10
Sco[["gbm"]] <- CreateDimReducObject(embeddings=out$V,key="GBM_")
Sco <- FindNeighbors(Sco,reduction = "gbm")
Sco <- FindClusters(Sco)
```


Plot the scores and color by the assigned clustering:

```{r}
plot_gbm(out, cluster=Sco$seurat_clusters)
```


Quantify the uncertainty in the low dimensional embedding:

```{r}
out <- get.se(out)

## Standard errors of V and U are now in the list
head(out$se_V) 
```

You can visualize the uncertainty with ellipses around the points

```{r}
library(ggforce)
plot_gbm(out, cluster=Sco$seurat_clusters, se=TRUE)
```


Now we evaluate cluster stability using the cluster confidence index. First we need to define a function that takes as input a set of simulated scores $\tilde{V}$ and returns a new clustering:

```{r}
cluster_fn <- function(V,Y) {
  Sco <- CreateSeuratObject(Y)
  colnames(V) <- 1:ncol(V)
  Sco[["gbm"]] <- CreateDimReducObject(embeddings=V,key="GBM_")
  Sco <- FindNeighbors(Sco,reduction = "gbm")
  Sco <- FindClusters(Sco)
  as.vector(Sco$seurat_clusters)
}
```

Now we can run the CCI function. Here we set `reps=10` to make it fast but `reps=100` (or higher) is recommended on real analyses. 

```{r, echo=T, eval=F}
cci <- CCI(out,cluster.orig=Sco$seurat_clusters, reps=10, cluster.fn = cluster_fn, Y=Y)
```

```{r, include=F}
cci <- CCI(out,cluster.orig=Sco$seurat_clusters, reps=10, cluster.fn = cluster_fn, Y=Y)
```


```{r}
library(pheatmap)
pheatmap(cci$H.table,legend=TRUE, color=colorRampPalette(c("white","red"))(100),
        breaks=seq(0,1,by=0.01),
        rownames=TRUE,
        colnames=TRUE)

#Just the diagonal
cci$cci.plot
```


The heatmap shows there is significant overlap between the clusters. This makes sense because the data was simulated to have no latent variability. 



